### 2022-12-06 paper.Rmd
discussion

Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity. In this paper, I presented the k-plus criterion, which focuses on the between-group similarity aspect of anticlustering. The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, higher order moments, and even covariances. Depending on the application, a different subset of these objectives can be included in the anticlustering process. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time is aggregated, especially in small data sets. In practice, users can try out different objectives and see which specification fulfils their needs. A reasonable approach is to start "greedy", e.g. by specifying that multiple distribution moments should equalized. If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variances). Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering. Moreover, using the open source R package `anticlust`, the application of k-plus anticlustering is easily accessible and free.

@@

The present paper focused on a new criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion. Instead of applying a specialized multicriterion optimization algorithm, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective. More sophisticated approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus criteria. @brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications. Their study focused on approximating the pareto set of efficient solutions that optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity. Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.

This paper's focus on the k-plus objective only instead of optimization algorithms followed two reasonings. First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications. Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it. To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective is optimized using the same algorithm, instead of using specialized optimization schemes tailored to each individual criterion. It was of particular interest to compare the k-plus objective with the common diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.

>>>>

Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity. In this paper, I presented the k-plus criterion, which focuses on maximizing between-group similarity. K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and even covariances. Depending on the application and the needs of a user, a different subset of these objectives can be included in the anticlustering process. The k-plus criterion is an extension of the classical k-means criterion, which only represents how similar groups are with regard to their means. Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by merely augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggravated in small data sets. In practice, users can try out different objectives and see which specification fulfils their needs. A reasonable approach is to start "greedy", e.g. by specifying that multiple distribution moments should equalized. If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variances). Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering. Moreover, using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.

@@

The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion. Instead of applying a specialized multicriterion optimization algorithm, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective. The same approach was used independent of the number of objectives that were included as part of the anticlustering process. More sophisticated algorithmic approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives. @brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications. Their study focused on approximating the pareto set of efficient solutions that optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity. Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.

This paper's focus on the k-plus objective only instead of optimization algorithms followed two reasonings. First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications. Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it. To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective is optimized using the same algorithm, instead of using specialized optimization schemes tailored to each individual criterion. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
